Testing Charts: viewers' perceptual accuracy in surveys

### 2023-07-23 SDSS-2023/testing-charts-accuracy.qmd
census estimates added to demographics

```{r}
#| label: fig-tasks
#| fig-cap: "Each of the visual stimuli presented to viewers. In each bar chart, two pieces are marked. The larger piece in each chart are pieces B and D. "
#| fig-subcap:
#| - "Vertical stacked bar, unaligned"
#| - "Horizontal stacked bar, unaligned"
#| - "Wide horizontal stacked bar, unaligned"
#| - "Vertical stacked bar, aligned"
#| - "Horizontal stacked bar, aligned"
#| - "Wide horizontal stacked bar, aligned"
#| layout-nrow: 2
#| out-width: 90%
knitr::include_graphics("images/vertical-ab.PNG")
knitr::include_graphics("images/round3/horizontal-ab.PNG")
knitr::include_graphics("images/round3/hwide-ab.PNG")
knitr::include_graphics("images/vertical-cd.PNG")
knitr::include_graphics("images/round3/horizontal-cd.PNG")
knitr::include_graphics("images/round3/hwide-cd.PNG")
```

@@

Task 1 was presented to viewers in one of two rounds (Rounds 1-2), with four color scheme variations each presented to 50% of respondents in each round. 
There were no significant differences in respondent accuracy or behavior across color schemes, and we combine them here into one set of stimulus responses to compare structural differences.
\kr{Should we put the color variations and a test for significance in accuracy/a visual in the appendix/supplementary materials? } \hh{we could do that, but I'm not sure that we need to mention color at all.}

@@

The visual mappings in the second task, shown in \autoref{fig-tasks}(b, e), are identical to the first task, but the axes of the chart are rotated so that the stacked bar is represented in a horizontal format.
This represents a structural change in how the data are presented to the viewer while preserving the pixel size of the elements viewers must compare.
Tasks 2 and 3 were asked within the same survey round (Round 3) and the full sample was randomly split among respondents, with 50% of participants seeing the Task 2 stimuli and 50% of participants seeing the Task 3 stimuli.

@@

Participants were recruited as part of NORC's AmeriSpeak panel, which utilizes a probability-based sampling methodology and samples U.S. households from NORC's National Sample Frame that provides coverage of over 97% of U.S. households.
The current panel size is 54,001 panel members aged 13 and over residing in over 43,000 households @dennis2019technical.
\hh{I don't think I've got the right citation} Each test was conducted using the AmeriSpeak Omnibus survey, which runs biweekly and samples around 1,000 U.S. adults to answer questions on a variety of topics. Our tests require visual inspection of an image; for this reason, our survey questions were only presented to web-based panelists and not panelists who respond via phone interviews.

@@

When combining responses from different rounds of the Omnibus survey, we calibrate weights in each round to the total population  across all rounds in the resulting model.

\kr{Do we need to update this description to just say how we did it for combining across all the rounds represented?}

We *combine* (rather than cumulate) $\ell \ge 2$ surveys $S_1$, $S_2$, ... $S_\ell$ as described in @omuircheartaighCombiningSamplesVs2002, by multiplying weights in $S_i$ by $\lambda_i$, for $1 \le i \le \ell$. $\lambda_i \in [0,1]$ with $\sum_i \lambda_i = 1$ is given as
$$
\lambda_i = \frac{n_i/d_i}{\sum_{j=1}^{\ell}n_j/d_j},
$$ where $n_i$ is the nominal sample of survey $S_i$ and $d_i$ are the design effects for the estimators.
Here, $d_i$ are estimated as 
$$
d_i = 1 + CV(w \in S_i)^2
$$ 
where $CV$ is the coefficient of variation of the weights $w$ within each sample, and is estimated as in @kishSurveySampling1965 :

$$ 
CV(w \in S) = \frac{\widehat{Var(w)}}{\bar{w}^2}.
$$

@@

The data for this paper were collected in several rounds as part of the NORC Omnibus, and the resulting number of participants, effective sample sizes, and $\lambda$ values are shown in \autoref{tbl-rounds}.

@@

| Name    | Date       |      \# Participants |                effective sample size |         Sum of weights $\sum_i w_i$ | $\lambda_i$ |
|---------------|---------------|--------------:|--------------:|--------------:|--------------:|
| Round 1 | April 2022 | `r nrow(round1_adj)` | `r round(neff(round1_adj$weight),1)` | `r round(sum(round1_adj$weight),1)` | $\lambda_1$  =  `r round(lambdas[1],3)` |
| Round 2 | May 2022   | `r nrow(round2_adj)` | `r round(neff(round2_adj$weight),1)` | `r round(sum(round2_adj$weight),1)` | $\lambda_2$  = `r sprintf("%.3f",lambdas[2])` |
| Round 3 | Jun 2022   | `r nrow(round3_adj)` | `r round(neff(round3_adj$weight),1)` | `r round(sum(round3_adj$weight),1)` | $\lambda_3$  = `r round(lambdas[3],3)` |
| **Total**        |    ---        |    `r nrow(round_123)`   |  `r round(neff(round_123$weight),1)`   |    |   |

@@

```{r}
#| label: fig-rounds-demographics
#| fig-cap: "Demographics of respondents"
#| fig-height: 7
#| fig-width: 12

@@

gender <- round_123 %>% 
  ggplot(aes(x = gender, weight = weight_pct_round)) + geom_bar() + 
  ylab("Percent") + 
  ggtitle("Gender") +
  theme(axis.text.x = element_text(angle=30, hjust = 1)) +
  xlab("") +
  facet_grid(.~chart)

age <- round_123 %>% 
  ggplot(aes(x = age4, weight = weight_pct_round)) + geom_bar() + 
  ylab("Percent") + 
  ggtitle("Age ") +
  theme(axis.text.x = element_text(angle=30, hjust = 1)) +
  xlab("") + ylim(c(0,35)) +
  facet_grid(.~chart)

education <- round_123 %>% 
  ggplot(aes(x = educ5, weight = weight_pct_round)) + geom_bar() + 
  ylab("Percent") + 
  ggtitle("Education levels") +
  theme(axis.text.x = element_text(angle=30, hjust = 1)) +
  xlab("") + ylim(c(0,35)) +
  facet_grid(.~chart)

>>>>

<!--
```{r eval=FALSE}
#| label: fig-tasks-old
#| fig-cap: "Each of the visual stimuli presented to viewers. In each bar chart, two pieces are marked. The larger piece in each chart are pieces B and D. "
#| fig-subcap:
#| - "Vertical, unaligned"
#| - "Horizontal, unaligned"
#| - "Wide horizontal, unaligned"
#| - "Vertical, aligned"
#| - "Horizontal, aligned"
#| - "Wide horizontal, aligned"
#| layout-nrow: 3
#| out-width: 90%
cat("Task 1")
cat("Task 2")
cat("Task 3")
knitr::include_graphics("images/vertical-ab.PNG")
knitr::include_graphics("images/round3/horizontal-ab.PNG")
knitr::include_graphics("images/round3/hwide-ab.PNG")
knitr::include_graphics("images/vertical-cd.PNG")
knitr::include_graphics("images/round3/horizontal-cd.PNG")
knitr::include_graphics("images/round3/hwide-cd.PNG")
```
-->

```{r}
#| label: fig-tasks
#| fig-cap: "Each of the visual stimuli presented to viewers. In each bar chart, two pieces are marked. The larger piece in each chart are pieces B and D. "
#| out-width: 90%
knitr::include_graphics("images/stimuli-overview.png")
```

@@

Task 1 was presented to viewers in two rounds (Rounds 1-2), with four color scheme variations each presented to 50% of respondents in each round. 
There were no significant differences in respondent accuracy or behavior across color schemes, and we combine them here into one set of stimulus responses to focus on the comparison of structural differences.
\kr{Should we put the color variations and a test for significance in accuracy/a visual in the appendix/supplementary materials? } \hh{we could do that, but I'm not sure that we need to mention color at all.}

@@

The visual mappings in the second task, shown in \autoref{fig-tasks}(b, e), are identical to the first task, but the axes of the chart are rotated so that the stacked bars are represented in a horizontal format.
This represents a structural change in how the data are presented to the viewer while preserving the pixel size of the elements viewers must compare.
Tasks 2 and 3 were asked within the same survey round (Round 3) and the full sample was randomly split among respondents, with 50% of participants seeing the Task 2 stimuli and 50% of participants seeing the Task 3 stimuli.

@@

Participants were recruited as part of NORC's AmeriSpeak panel, which utilizes a probability-based sampling methodology and samples U.S. households from NORC's National Sample Frame that provides coverage of over 97% of U.S. households.
The current panel size is 54,001 panel members aged 13 and over residing in over 43,000 households [@dennis2019technical].
\hh{I don't think I've got the right citation} Each test was conducted using the AmeriSpeak Omnibus survey, which runs biweekly and samples around 1,000 U.S. adults to answer questions on a variety of topics. Our tests require visual inspection of an image; for this reason, our survey questions were only presented to web-based panelists and not panelists who respond via phone interviews.

@@

When combining responses from different surveys, weights need to be calibrated to the total population, so that their order after combining still reflects their relative importance.

<!--
\kr{Do we need to update this description to just say how we did it for combining across all the rounds represented?}\hh{done} -->

We *combine* (rather than cumulate) a set of $\ell$ surveys $S_1$, $S_2$, ... $S_\ell$ (with $ \ell \ge 2$ ) as described in @omuircheartaighCombiningSamplesVs2002, by multiplying weights in $S_i$ by $\lambda_i$, for $1 \le i \le \ell$. $\lambda_i \in [0,1]$ with $\sum_i \lambda_i = 1$ is given as
$$
\lambda_i = \frac{n_i/d_i}{\sum_{j=1}^{\ell}n_j/d_j},
$$ where $n_i$ is the nominal sample of survey $S_i$ and $d_i$ are the design effects for the estimators.
Here, $d_i$ are estimated as 
$$
d_i = 1 + CV(w \in S_i)^2
$$ 
where $CV$ is the coefficient of variation of the weights $w$ within each sample, and is estimated as in @kishSurveySampling1965 :
$$ 
CV(w \in S) = \frac{\widehat{Var(w)}}{\bar{w}^2}.
$$

@@

The data for this paper were collected in several rounds as part of the NORC Omnibus. The resulting number of participants, effective sample sizes, and $\lambda$ values are shown in \autoref{tbl-rounds}.

@@

| Name    | Date       |      \# Participants |                effective sample size |         Sum of weights $\sum_i w_i$ | $\lambda_i$ |
|---------------|---------------|--------------:|--------------:|--------------:|--------------:|
| Round 1 | April 2022 | `r nrow(round1_adj)` | `r round(neff(round1_adj$weight),1)` | `r round(sum(round1_adj$weight),1)` | $\lambda_1$  =  `r round(lambdas[1],3)` |
| Round 2 | May 2022   | `r nrow(round2_adj)` | `r round(neff(round2_adj$weight),1)` | `r round(sum(round2_adj$weight),1)` | $\lambda_2$  = `r sprintf("%.3f",lambdas[2])` |
| Round 3 | June 2022   | `r nrow(round3_adj)` | `r round(neff(round3_adj$weight),1)` | `r round(sum(round3_adj$weight),1)` | $\lambda_3$  = `r round(lambdas[3],3)` |
| **Total**        |    ---        |    `r nrow(round_123)`   |  `r round(neff(round_123$weight),1)`   |    |   |

@@

```{r}
#| label: fig-rounds-demographics
#| fig-cap: "Demographics of respondents in each of the tasks. The dots show percentages of the national demographics as estimated by the American Community Survery (ACS) 2021."
#| fig-height: 7
#| fig-width: 12

@@

# from ACS 2001
# https://data.census.gov/table?q=sex+of+adults+age+greater+than+18+across+the+US&tid=ACSDP1Y2021.DP05	
acs_gender <- data.frame(gender = c("Male", "Female"), 
                         estimate = c(126690177, 131728290), moe=c(22354, 18838),
                         percent = c(49, 51), moe_percent = 0.1)

gender <- round_123 %>% 
  ggplot(aes(x = gender)) + 
  geom_bar(aes(weight = weight_pct_round), fill="grey", colour = "grey30", linewidth=0.5) + 
  ylab("Percent") + 
  ggtitle("Gender") +
  theme(axis.text.x = element_text(angle=30, hjust = 1)) +
  xlab("") +
  facet_grid(.~chart) + 
  geom_point(aes(x = gender, y = percent), data = acs_gender)

acs_age <- data.frame(age4 = c("18-29", "30-44", "45-59", "60+"), 
                      estimate = c(52325456, 66712246, 61814869, 77565896))
acs_age <- acs_age %>% mutate(percent = estimate/sum(estimate)*100)

age <- round_123 %>% 
  ggplot(aes(x = age4)) + 
  geom_bar(aes(weight = weight_pct_round), fill="grey", colour = "grey30", linewidth=0.5) + 
  ylab("Percent") + 
  ggtitle("Age ") +
  theme(axis.text.x = element_text(angle=30, hjust = 1)) +
  xlab("") + ylim(c(0,35)) +
  facet_grid(.~chart) +
  geom_point(aes(x = age4, y = percent), data = acs_age)

acs_educ5 <- data.frame(educ5 = c("Less than HS", "HS or equivalent", "Some college",  "Bachelor", "Post graduate"), 
                      estimate = c(28687047, 69610758, 77707408, 50031690, 29454503))
acs_educ5 <- acs_educ5 %>% mutate(percent = estimate/sum(estimate)*100)

education <- round_123 %>% 
  ggplot(aes(x = educ5)) + 
  geom_bar(aes(weight = weight_pct_round), fill="grey", colour = "grey30", linewidth=0.5) + 
  ylab("Percent") + 
  ggtitle("Education levels") +
  theme(axis.text.x = element_text(angle=30, hjust = 1)) +
  xlab("") + ylim(c(0,35)) +
  facet_grid(.~chart) +
  geom_point(aes(y = percent), data = acs_educ5)
